---
title: "Working with the seedclim community database"
author: "Richard Telford"
date: "4/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This vignette demonstrates how to use the seedclim community database.
This is useful for analysing the community data, but also for accessing the taxon or site tables used by other data types.

The code in this vignette can be copied and modified to form the data importing section of your own code.

## Downloading the database

You only need to do this once, unless you want to check for an updated version of the data.


```{r download, eval = TRUE}
library("dataDownloader")
library("here")

# get_file will only download the file if it is missing or it needs updating
node <- "npfa9"
path <- "community/data/"
file <- "seedclim.2020.4.15.zip"
get_file(node = node, 
         file = file,
         path = here(path),
         remote_path = "3_Community_data")

#unzip
unzip(here(path, file), exdir = here(path))
```

## Connecting to the database

The seedclim community database is an SQLite database. 
You can connect to the database with `DBI::dbConnect`

```{r connect, message = FALSE}
#load packages
library("tidyverse")
library("DBI")# also needs RSQLite installed
#devtools::install_github("cynkra/dm") need dev version
library("dm")
#make database connection
con <- dbConnect(RSQLite::SQLite(), here(path, "seedclim.sqlite"))
```


## Database tables

```{r tables}
tables <- dbListTables(conn = con)
tables
```

## Data model

```{r dm, fig.cap = "Relationship between tables in the database, showing primary and foreign", echo = FALSE}

dm_from_src(con, learn_keys = FALSE) %>%
  dm_add_pk(sites, siteID) %>% 
  dm_add_fk(blocks, siteID, sites) %>% 
  dm_add_fk(site_attributes, siteID, sites) %>% 
  dm_add_pk(site_attributes, c(siteID, attribute)) %>% 
  dm_add_pk(blocks, blockID) %>% 
  dm_add_fk(plots, blockID, blocks) %>% 
  dm_add_pk(plots, plotID) %>% 
  dm_add_fk(turfs, originPlotID, plots) %>% 
  dm_add_fk(turfs, destinationPlotID, plots) %>% 
  
  dm_add_pk(turfs, turfID) %>% 
  dm_add_fk(turf_community, turfID, turfs) %>% 
  dm_add_fk(subturf_community, turfID, turfs) %>% 
  dm_add_fk(turf_environment, turfID, turfs) %>% 
  dm_add_fk(subturf_environment, turfID, turfs) %>%
    
  dm_add_pk(taxon, species) %>% 
  dm_add_fk(turf_community, species, taxon) %>% 
  dm_add_fk(subturf_community, species, taxon) %>%
  dm_add_pk(turf_community, c("turfID", "species", "year")) %>%  
  dm_add_pk(turf_environment, c("turfID", "year")) %>% 
  dm_add_pk(subturf_community, c("turfID", "subturf", "species", "year")) %>%  
  dm_add_pk(subturf_environment, c("turfID", "subturf", "year")) %>% 
  
  dm_add_fk(species_attributes, species, taxon) %>% 
  dm_add_pk(species_attributes, c(species, attribute)) %>% 
  dm_add_pk(attributes, attribute) %>% 
  dm_add_fk(site_attributes, attribute, attributes) %>% 
  dm_add_fk(species_attributes, attribute, attributes) %>% 
  
  dm_draw(view_type = "keys_only")
  

```


## Importing data

Data can be imported with `dplyr::tbl`. 
The first argument is the database connection, the second is the name of the table 
Then the data can be manipulated with the full range of tidyverse functions.
Remember to use `dplyr::collect` to collect all the data rather than just the first few rows.

```{r importing}
taxa <- tbl(src = con, "taxon") %>% 
  collect

```

You can also use a SQL statement in wrapped in `sql()`, should you really want to.

```{r importing-sql}
tbl(src = con, sql("SELECT * FROM taxon"))
```

## Importing cover data

To import the cover data, we need to import data from several tables and join them.
The transplanted turfs have an origin site and a destination site (connected via the plots and blocks), so we need to join site_block_plot to the turf data twice, once with `originPlotID` and once with `destinationPlotID`.


```{r cover}
sites_blocks_plots <- tbl(con, "sites") %>% 
  inner_join(tbl(con, "blocks"), by = "siteID") %>% 
  inner_join(tbl(con, "plots"), by = "blockID") %>% 
  select(siteID, norwegian_name, annual_precipitation_gridded, summer_temperature_gridded,
         temperature_level, precipitation_level, blockID, plotID)

cover_thin <- sites_blocks_plots %>% 
  inner_join(tbl(con, "turfs"), by = c("plotID" = "originPlotID")) %>% 
  inner_join(sites_blocks_plots, by = c("destinationPlotID" = "plotID"), 
             suffix = c("_origin", "_dest")) %>% 
  inner_join(tbl(con, "turf_community"), by = "turfID") %>% 
  select(-plotID, -destinationPlotID, -cover_raw) %>%
  collect()

cover_thin
```

Now we need to filter the data used for each experiment.

- The turf transplant experiment turfs have an entry in the `TTtreat` field.

The other experiments are weird - will be fixed

<!-- - The grass removal experiment turfs have an entry in the `GRtreat` field. -->
<!-- - The recruit tag experiment turfs have an entry in the `RTtreat` field. -->

Control turfs can be the control for multiple experiments.

```{r filter-TTtreat}
cover_thin_ttreat <- cover_thin %>% 
  filter(
    TTtreat != "", # only TTtreat. Also remove NA values from this field
    year != 2010  # no TTtreat data for 2010, just other experiments sharing the turfs
  )
```

## Importing subturf data


```{r subturf}
#TODO
```


Turf and subturf environment data can be imported in a similar way, or joined to, e.g., `cover_thin`.
