---
title: "Gridded climate data"
author: "Aud Halbritter"
date: "19.12.2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This vignette demonstrates how to use the seedclim gridded climate data.
This data is useful as covariable in many analysis.

The code in the vignette can be copied and modified to form the data importing section of your own code.


```{r loadStuff, echo = FALSE}

#Load packages
library("tidyverse")
library("lubridate")

# read in data to R
climate <- read_csv(file = "climate/data/GriddedDailyClimateData2009-2019.csv", col_names = TRUE)

```

##  Overview

The gridded climate data was provided by met.no (Ole Einar Tveito, 25.10.2016, 17.01.2017, 16.12.2019)
The data are daily climate data for the 12 SeedClim sites between 2009 and 2019 (will be updated).


The climate variables are  
 - Mean daily temperature (°C, temperature)  
 - Relative air moisture (%, rel_air_moisture)  
 - Mean wind (meter / second, wind)  
 - Mean cloud cover (in 8 parts, no dimension, cloud_cover)  
 - Precipitation (mm, precipitation)

## Raw data and processing

The raw data is available as .dat files and is stored on OSF (https://osf.io/npfa9/) as a zip file (Gridded_Climate_Data_2009-2019.zip) in the "8_Environmental_data/Raw_data" folder.
The code to process the raw data can be found in this GitHub repo (climate/ReadInGriddedData.R).


## Importing the data

The data are provided as a csv file on OSF (https://osf.io/npfa9/).

The data can be downloaded directly to are using the dataDownloader R package.

```{r osf-download, eval=FALSE, echo=TRUE} 

# install.packages("remotes")
#remotes::install_github("Between-the-Fjords/dataDownloader")
library("dataDownloader")

node <- "npfa9"
filename <- "GriddedDailyClimateData2009-2019.csv",
pathname <- "climate/data/"

# Download data from OSF
get_file(node = node,
         file = filename,
         path = here(pathname),
         remote_path = "8_Environmental_data")

```


The data can be read in using the following code.
The data are provided in a tidy format, containing date, siteID (corresponding with the rest of the SeedClim data), variable (climate variable) and value.

```{r data-import, eval=TRUE, echo=TRUE}

# read in data to R
climate <- read_csv(file = here(pathname, filename))
climate

```


## Calculate monthly and annual values

To calculate monthly and annual climate data this code can be used.

```{r calculations, eval=TRUE, echo=TRUE}
# Calculate monthly climate data
monthly_climate <- climate %>%
  mutate(date_month = dmy(paste0("15-",format(date, "%b.%Y")))) %>%
  group_by(date_month, variable, siteID) %>%
  summarise(n = n(), 
            value = mean(value), 
            sum = sum(value)) %>% 
  mutate(value = ifelse(variable == "precipitation", sum, value)) %>% 
  select(-n, -sum)


# Calculate annual climate data
annual_climate <- monthly_climate %>% 
  mutate(year = year(date_month)) %>%
  group_by(year, variable, siteID) %>%
  summarise(annual_value = mean(value))

```


## Plotting the data

### Temperature

```{r temp, eval=TRUE, echo=TRUE}
p <- climate %>% 
  pivot_wider(names_from = "variable", values_from = "value") %>% 
  mutate(siteID = factor(siteID, levels = c("Ulvehaugen", "Lavisdalen", "Gudmedalen", "Skjelingahaugen", "Alrust", "Hogsete", "Rambera", "Veskre", "Fauske", "Vikesland", "Arhelleren", "Ovstedalen", "Joasete", "Liahovden"))) %>%
  ggplot(aes(x = date, y = temperature, color = siteID)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ siteID, ncol = 4) + 
  theme(legend.position = "none") +
  ggtitle("Mean daily temperature °C")
p
```


RelAirMoisture

```{r moist, eval=TRUE, echo=TRUE}
p + aes(y = rel_air_moisture) +
  ggtitle("Relative air moisture %")
```


Wind
```{r wind, eval=TRUE, echo=TRUE}
p + aes(y = wind) +
  ggtitle("Wind meter per second")

```


CloudCover
```{r cloud, eval=TRUE, echo=TRUE}
p + aes(y = cloud_cover) +
  ggtitle("Cloud cover")
```


Precipitation
```{r prec, eval=TRUE, echo=TRUE}
p + aes(y = precipitation) +
  ggtitle("precipitation mm")

```



